Just Take the Bus: Analysis of Toronto Bike Theft from 2014-2023

### 2024-11-12 Paper/Article.Rmd
Code comments and fixes

---
title: "Just Take the Bus: Analysis of Toronto Bike Theft from 2022-2024"
subtitle: tbd
author:
- Rayan Awad Alim
- Samreen Gill
thanks: https://github.com/RayanAlim/Toronto_Bike_Thefts_Data_Analysis
date: "November 12, 2024"
abstract: This analysis examines Bike theft data obtained from the Tronrot Police Data Portal.
output: pdf_document
toc: true
number-sections: true
editor_options:
  markdown:
    wrap: 72
---

@@

knitr::opts_chunk$set(echo=FALSE, include=TRUE, eval=TRUE)
required_packages <- c(
  "lubridate", "sf", "tidyverse", "ggplot2", "maps", "dplyr", "sf", "mapdata", "ggspatial", "opendatatoronto", "here"
) 
for (p in required_packages) { # install if missing
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

library(ggplot2)
library(dplyr)
library(tidyverse)
library(sf)
library(mapdata)
library(ggspatial)
library(opendatatoronto)
library(here)
```

@@

@@

```{r}
#BY MONTH
bike_thefts %>%
  count(OCC_MONTH) %>%
  ggplot(aes(x = reorder(OCC_MONTH, n), y = n, fill = OCC_MONTH)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Thefts by Month of the Year",
       x = "Month", y = "Number of Thefts") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r}
# BY DAY OF WEEK
bike_thefts %>%
  count(OCC_DOW) %>%
  ggplot(aes(x = OCC_DOW, y = n, fill = OCC_DOW)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Thefts by Day of the Week",
       x = "Day of the Week", y = "Number of Thefts") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
```

```{r}
# heaatmap of thefts by day of week and hour

@@

```{r}
#TOP TEN 
bike_thefts %>%
  count(NEIGHBOURHOOD_158) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = reorder(NEIGHBOURHOOD_158, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Neighborhoods for Bike Thefts",
       x = "Neighborhood", y = "Number of Thefts") +
  theme_minimal()
```

```{r}
#| include: false
#| warning: false
#| message: false
#MAP TBD
#library(opendatatoronto)
```

@@

```{r}
# Barplot for bike types
ggplot(bike_thefts, aes(x = BIKE_TYPE)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Number of Bike Thefts by Bike Type", x = "Bike Type", y = "Count") +
  theme_minimal()
```

```{r}
# Box plot for bike cost distribution by status
ggplot(bike_thefts %>% filter(!is.na(BIKE_COST)), aes(x = STATUS, y = BIKE_COST)) +
  geom_boxplot() +
  labs(title = "Bike Cost Distribution (Recovered vs Stolen)", x = "Status", y = "Cost ($)") +
  theme_minimal()
```

@@

```{r}
#| include: FALSE
#| warning: false
#| message: false
# calcuate reporting delay
bike_thefts <- bike_thefts %>%
  mutate(reporting_delay = as.numeric(difftime(ymd_hms(REPORT_DATE), ymd_hms(OCC_DATE), units = "days")))

@@

```{r}
# map

# Count and filter top 10 neighborhoods with the highest thefts
top_neighborhoods <- bike_thefts %>%
  count(NEIGHBOURHOOD_158) %>%
  slice_max(n, n = 10)

# Treemap plot
treemap::treemap(
  top_neighborhoods,
  index = "NEIGHBOURHOOD_158",  # Using neighborhood names as the index
  vSize = "n",                  # Size by count of thefts
  vColor = "n",                 # Color by count of thefts
  palette = c("#2b5c8a", "goldenrod1", "#9e3d22"),
  type = "manual",
  align.labels = list(c("right", "top"), c("left", "bottom")),
  overlap.labels = 0.2,
  title = "Number of Bicycle Thefts by Neighborhood",
  fontsize.title = 18,
  title.legend = "Frequency",
  fontsize.labels = 10, 
)
```

>>>>

---
title: "Just Take the Bus: Analysis of Toronto Bike Theft from 2014-2023"
subtitle: tbd
author:
- Rayan Awad Alim
- Samreen Gill
thanks: https://github.com/RayanAlim/Toronto_Bike_Thefts_Data_Analysis
date: "November 12, 2024"
abstract: This analysis examines Bike theft data obtained from the Tronrot Police Data Portal.
output: pdf_document
toc: true
number-sections: true
editor_options:
  markdown:
    wrap: 72
---

@@

knitr::opts_chunk$set(echo=FALSE, include=TRUE, eval=TRUE)
required_packages <- c(
  "lubridate", "sf", "tidyverse", "ggplot2", "maps", "dplyr", "sf", "mapdata", "ggspatial", "opendatatoronto", "here", "treemap"
) 
for (p in required_packages) { # install if missing
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

library(ggplot2)
library(dplyr)
library(tidyverse)
library(sf)
library(mapdata)
library(ggspatial)
library(opendatatoronto)
library(here)
library(treemap)
```

@@

```{r}
# Samreen 
# bar charts of general statistics:
# # of bikes stole in 2021
# # of bikes stole in 2022
# # of bikes stole in 2023
# Total Loss of x $ M 
```

@@

```{r}
#BY MONTH- Rayan
# Note: Replace Legend title to "month" rather than variable name 
# Order it based on month 
bike_thefts %>%
  count(OCC_MONTH) %>%
  ggplot(aes(x = reorder(OCC_MONTH, n), y = n, fill = OCC_MONTH)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Thefts by Month of the Year",
       x = "Month", y = "Number of Thefts") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r}
# BY DAY OF WEEK- Rayan
# Note: Replace Legend title to "day" rather than varaible name 
# Order it based on day 
bike_thefts %>%
  count(OCC_DOW) %>%
  ggplot(aes(x = OCC_DOW, y = n, fill = OCC_DOW)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Thefts by Day of the Week",
       x = "Day of the Week", y = "Number of Thefts") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
```

```{r}
# heaatmap of thefts by day of week and hour
# flip the time from 0-24 top to bottom
# make "count" more meaningful 
#Organize it by day of the week - Rayan

@@

```{r}
#TOP TEN 
#Remove neighbour hood # from the name - Rayan
bike_thefts %>%
  count(NEIGHBOURHOOD_158) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = reorder(NEIGHBOURHOOD_158, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Neighborhoods for Bike Thefts",
       x = "Neighborhood", y = "Number of Thefts") +
  theme_minimal()
```
```{r}
#colour to be something categorical #https://r-graph-gallery.com/38-rcolorbrewers-palettes.html - Rayan
library(RColorBrewer)

bike_thefts %>%
  count(NEIGHBOURHOOD_158) %>%
  arrange(desc(n)) %>%  
  slice_max(n, n = 10) %>%
  ggplot(aes(x = "", y = n, fill = reorder(NEIGHBOURHOOD_158, -n))) +  
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0("(", round(n / sum(n) * 100, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "white",
            size = 4) +
  labs(title = "Top 10 Neighborhoods for Bike Thefts",
       fill = "Neighborhood") +
  scale_fill_brewer(palette = "set3") +
  theme_void() +  
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
#| include: false
#| warning: false
#| message: false
#MAP TBD
#library(opendatatoronto)
# MAE THE MAP WORK! - Rayan
```

@@

```{r}
# Barplot for bike types
# Add a legend with full name of bike type
ggplot(bike_thefts, aes(x = BIKE_TYPE)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Number of Bike Thefts by Bike Type", x = "Bike Type", y = "Count") +
  theme_minimal()
```
### Cost

```{r}
# Pie Chart of recovery vs stolen
```

@@

```{r}
# Pie Chart of recovery vs stolen
```

```{r}
#| include: FALSE
#| warning: false
#| message: false
# calcuate reporting delay
# reporting vs theft and status of revovey
bike_thefts <- bike_thefts %>%
  mutate(reporting_delay = as.numeric(difftime(ymd_hms(REPORT_DATE), ymd_hms(OCC_DATE), units = "days")))

@@
